---
output: github_document
bibliography: inst/REFERENCES.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "75%"
)
```

# GGMnonreg

 
[![CircleCI build status](https://circleci.com/gh/donaldRwilliams/GGMnonreg.svg?style=svg)](https://circleci.com/gh/donaldRwilliams/GGMnonreg)


The goal of **GGMnonreg** is to estimate non-regularized Gaussian graphical models. Note that 
the title is a bit of a misnomer, in that Ising and mixed graphical models are also supported.

Graphical modeling is quite common in fields with *wide* data, that is, when there are more 
variables than observations. Accordingly, many regularization-based approaches have been developed for those kinds of data. There are key drawbacks of regularization when the goal is inference, 
including, but not limited to, the fact that obtaining a valid measure of parameter uncertainty is very (very) difficult.

More recently, graphical modeling has emerged in psychology [@Epskamp2018ggm], where the data 
is typically long or low-dimensional [*p* < *n*; @williams2019nonregularized; @williams_rethinking]. The primary purpose of  **GGMnonreg** is to provide methods specifically for low-dimensional data 
(e.g., those common to psychopathology networks).

## Supported Models
* Gaussian graphical model. The following data types are supported.
  + Gaussian 
  + Ordinal
  + Binary
* Ising model [@marsman_2018]
* Mixed graphical model

## Additional methods
The following are also included

* Expected network replicability [@williams2020learning]
* Compare Gaussian graphical models
* Measure of parameter uncertainty [@williams2019nonregularized]
* Edge inclusion "probabilities"
* Network visualization with **ggplot2** [@ggplotpackage]
* Constrained precision matrix (the network, given an assumed graph)
* Predictability (variance explained)

## Installation
You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("donaldRwilliams/GGMnonreg")
```

## Ising
An Ising model is fitted with the following
```{r}
library(GGMnonreg)

# make binary
Y <- ifelse(ptsd[,1:5] == 0, 0, 1)

# fit model
fit <- ising_search(Y, IC = "BIC", 
                    progress = FALSE)

fit
```

Note the same code, more or less, is also used for GGMs and mixed graphical models.

## Predictability
It is common to compute predictability, or variance explained, for each node in the network.
An advantage of **GGMnonreg** is that a measure of uncertainty is also provided.

```{r}
# data
Y <- na.omit(bfi[,1:5])

# fit model
fit <- ggm_inference(Y, boot = FALSE)

# predictability
predictability(fit)
```


## Parameter Uncertainty
Confidence intervals for each relation are obtained with
```{r}
# data
Y <- na.omit(bfi[,1:5])

# fit model
fit <- ggm_inference(Y, boot = TRUE, 
                     method = "spearman", 
                     B = 100, progress = FALSE)

confint(fit)
```
These can then be plotted with, say, **ggplot2** (let to the user).

## Edge Inclusion
When mining data, or performing an automatic search, it is difficult to make inference on the
network parameters (e.g., confidence are not easily computed). To summarize data mining,
**GGMnonreg** provides edge inclusion "probabilities" (proportion bootstrap samples for 
which each relation was detected).

```{r}
# data
Y <- na.omit(bfi[,1:5])

# fit model
fit <-  eip(Y, method = "spearman", 
            B  = 100, progress = FALSE)

fit
```
Note in all cases, the provided estimates correspond to the upper-triangular elements
of the network.

## Expected Network Replicability
**GGMnonreg** allows for computing expected network replicability (ENR), i.e., the number of 
effects that will be detected in any number of replications. This is an analytic solution.

The first step is defining a true network
```{r}
# first make the true network
cors <- cor(GGMnonreg::ptsd)

# inverse
inv <- solve(cors)

# partials
pcors <-  -cov2cor(inv)

# set values to zero
pcors <- ifelse(abs(pcors) < 0.05, 0, pcors)
```

Then obtain ENR
```{r}
fit_enr <- enr(net = pcors, n = 500, replications = 2)

fit_enr
```
Note this is inherently frequentist. As such, over the long run, 45 % of the edges will be replicated on average. Then we can further infer that, in hypothetical replication attempts, more than half of the edges
will be replicated only 5 % of the time.

ENR can also be plotted
```{r}
plot(fit_enr)
```

### Intution
Here is the basic idea of ENR
```{r}
# location of edges
index <- which(pcors[upper.tri(diag(20))] != 0)

# convert network into correlation matrix
diag(pcors) <- 1
cors_new <- corpcor::pcor2cor(pcors)

# replicated edges
R <- NA

# increase 1000 to, say, 5,000
for(i in 1:1000){

  # two replications
  Y1 <- MASS::mvrnorm(500, rep(0, 20), cors_new)
  Y2 <- MASS::mvrnorm(500, rep(0, 20), cors_new)

  # estimate network 1
  fit1 <- ggm_inference(Y1, boot = FALSE)

  # estimate network 2
  fit2 <- ggm_inference(Y2, boot = FALSE)

  # number of replicated edges (detected in both networks)
  R[i] <- sum(
    rowSums(
      cbind(fit1$adj[upper.tri(diag(20))][index],
            fit2$adj[upper.tri(diag(20))][index])
    ) == 2)
}
```
Notice that replication of two networks is being assessed over the long run. In other words,
if we draw two random samples, what is the expected replicability.

Compare analytic to simulation
```{r}
# combine simulation and analytic
cbind.data.frame(
  data.frame(simulation = sapply(seq(0, 0.9, 0.1), function(x) {
    mean(R > round(length(index) * x) )
  })),
  data.frame(analytic = round(fit_enr$cdf, 3))
)

# average replicability (simulation)
mean(R / length(index))

# average replicability (analytic)
fit_enr$ave_pwr
```

ENR works with any correlation, assuming there is an estimate of the standard error.

## Network plot
```{r, message=FALSE}
# data
Y <- ptsd

# estimate graph
fit <- ggm_inference(Y, boot = FALSE)

# get info for plotting
plot(get_graph(fit), edge_magnify = 5)
```

## References
